---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
meteorite_data <- read_csv("meteorite_landings.csv") %>%clean_names
```

```{r}
view(meteorite_data)
```

```{r}
dim(meteorite_data)
```

```{r}
glimpse(meteorite_data)
```

```{r}
meteorite_data %>% 
  distinct(fall)
```


```{r}
#Could have done it this way but janitor seemed to clean everything up
#renamed_meteorite_data <-
meteorite_data %>%
  rename("mass_grams" = `mass (g)`) %>% 
  rename("fell_or_found" = fall) %>% 
  rename("year_of_discovery" = year) %>% 
  rename("geo_location" = GeoLocation)

```

```{r}
#However, did want to change the name of the year and the fell or found as above
renamed_meteorite_data <-
  meteorite_data %>%
  rename("fell_or_found" = fall) %>% 
  rename("year_of_discovery" = year) 
  
renamed_meteorite_data
```


Split in column GeoLocation into latitude and longitude, the new latitude and longitude columns should be numeric.

```{r}
renamed_meteorite_data %>%
sample_n(10) %>% 
  select(geo_location)
```


```{r}
q1_1_3_split <-
renamed_meteorite_data %>%
  
  separate(geo_location, c("latitude", "longitud"), sep=",")
 
q1_1_3_split

```

```{r}
q1_1_3_split <-
q1_1_3_split %>% 
  mutate(latitude = str_replace(latitude, "[(]", "")) %>%
  mutate(longitud = str_replace(longitud, "[)]", "")) 

q1_1_3_split

```

```{r}
q1_1_3_split <-
q1_1_3_split %>%
  mutate(latitude = as.numeric(latitude)) %>% 
  mutate(longitud = as.numeric(longitud))

q1_1_3_split
```

Q1.1.4 Replace any missing values in latitude and longitude with zeros.

```{r}
q1_1_3_split %>% 
  summarise(geo_nas = sum(is.na (c(latitude, longitud))))
```

```{r}
q1_1_4_na_clean <-
  q1_1_3_split %>%
  mutate(latitude = coalesce(latitude, 0, na.rm = TRUE)) %>%
  mutate(longitud = coalesce(longitud, 0, na.rm = TRUE))

q1_1_4_na_clean

```

```{r}
q1_1_4_na_clean %>% 
  summarise(geo_nas = sum(is.na (c(latitude, longitud))))
```

Q1.1.5 Remove meteorites less than 1000g in weight from the data.

```{r}
q1_1_5_remove_small_meteorites <-
q1_1_4_na_clean %>%
  filter(mass_g >= 1000)

q1_1_5_remove_small_meteorites
```

Q1.1.6 Order the data by the year of discovery.

```{r}
q1_1_6_arrange_by_year <-
q1_1_5_remove_small_meteorites %>%
  arrange(year_of_discovery)

q1_1_6_arrange_by_year
```

We would also like you to include assertive programming to make sure that:

1. The data has the variable names we expect (“id”, “name”, “mass (g)”, “fall”, “year”, “GeoLocation”).
2. Latitude and longitude are valid values. (Latitude between -90 and 90, longitude between -180 and 180).

```{r}
library(assertr)
```

```{r}
check_lat_long <- function(lat_long_num){
  lat_long_num %>% 
    verify(latitude >= -90 & latitude <= 90) %>%
    verify(longitud >= -180 & longitud <= 180)
  return(list(lat_long_num))
}

```

```{r}
check_lat_long(q1_1_6_arrange_by_year)
```

```{r}
q1_1_6_arrange_by_year %>%
sample_n(10) %>%
  select(name, latitude)
```

```{r}
cleaned_meteorite_data <- q1_1_6_arrange_by_year
```

```{r}
cleaned_meteorite_data %>% 
write_csv(file = "cleaned_meteorite_data.csv")
```

